---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
#DEATHS CODE
library(dplyr)  # for data manipulation
library(stringr)  # for string operations

# Function to convert "YES", "NO", and "" to the desired format
convert_status <- function(x) {
  ifelse(x == "YES", "YES", ifelse(x == "NO", "NO", ""))
}

library(tidyr)  # for data reshaping
library(readr)

# Reshape Death columns into Time and Death columns
deaths <- av %>%
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") %>%
  mutate(Time = parse_number(Time),
         Death = convert_status(Death)) %>%
  select(Time, Death)

# Calculate the average number of deaths per Avenger
average_deaths <- deaths %>%
  filter(Death == "YES") %>%
  summarise(avg_deaths = mean(Time, na.rm = TRUE))

# Print the average number of deaths
print(average_deaths)

```

```{r}
#RETURNS CODE
library(dplyr)  # for data manipulation
library(stringr)  # for string operations

# Function to convert "YES", "NO", and "" to the desired format
convert_statusr <- function(x) {
  ifelse(x == "YES", "YES", ifelse(x == "NO", "NO", ""))
}

library(tidyr)  # for data reshaping
library(readr)

# Reshape Return columns into Time and Return columns
returns <- av %>%
  pivot_longer(cols = starts_with("Return"),
               names_to = "Time",
               values_to = "Return") %>%
  mutate(Time = parse_number(Time),
         Death = convert_statusr(Return)) %>%
  select(Time, Return)

# Calculate the average number of returns per Avenger
average_returns <- returns %>%
  filter(Return == "YES") %>%
  summarise(avg_returns = mean(Time, na.rm = TRUE))

# Print the average number of returns
print(average_returns)
```


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement
> Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team. 

### Connor's Code
```{r}
# Count occurrences of "YES" in Death column for each period in Time column
death_counts <- deaths %>%
  group_by(Time) %>%
  count(Death) %>%
  filter(Death == "YES") %>%
  rename(Deaths_YES = n) %>%
  select(Time, Deaths_YES)

# Print the resulting data frame
print(death_counts)
```
### Connor's answer

***As we can see, the number of avengers who died at least once is 69. This proves that the statement that 69 out of 173 avengers have died is true.***

### Melania's Code & Answer
***"What’s more, if you’re a fan of the MCU, nobody is safe. Of the nine Avengers we see on screen — Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once in the course of their time Avenging in the comics. In fact, Hawkeye died twice!"***
```{r}
# Filter to only include the nine Avengers mentioned in the quote
avengers <- av %>%
  filter(Name.Alias %in% c('Anthony Edward "Tony" Stark', "Robert Bruce Banner", "Steven Rogers", "Thor Odinson", "Clinton Francis Barton", "Natalia Alianovna Romanova", "Wanda Maximoff", "Pietro Maximoff", "Victor Shade (alias)"))
avengers

# Check if each avenger has at least one "YES" in deaths
avengers <- avengers %>%
  mutate(has_died = rowSums(select(., starts_with("Death")) == "YES") > 0)
avengers$has_died
```
***Each of the nine Avengers mentioned has died at least once in the course of their time.***

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
